Semi-Parametric Manifold Clustering

### 2021-11-17 curve-fitting-methods/curve-fitting-scratch.Rmd
still working on typing up new material

\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{Tr}}
\newcommand{\blockdiag}{\mathrm{blockdiag}}
\newcommand{\indep}{\stackrel{\mathrm{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
\newcommand{\Bernoulli}{\mathrm{Bernoulli}}
\newcommand{\Betadist}{\mathrm{Beta}}
\newcommand{\Uniform}{\mathrm{Uniform}}
\newcommand{\BG}{\mathrm{BernoulliGraph}}
\newcommand{\Categorical}{\mathrm{Categorical}}
\newcommand{\Multinomial}{\mathrm{Multinomial}}
\newcommand{\RDPG}{\mathrm{RDPG}}
\newcommand{\GRDPG}{\mathrm{GRDPG}}
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\theoremstyle{example}
\newtheorem{example}{Example}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\as}{\stackrel{\mathrm{a.s.}}{\to}}

@@

>>>>

\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{Tr}}
\newcommand{\blockdiag}{\mathrm{blockdiag}}
\newcommand{\indep}{\stackrel{\mathrm{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
\newcommand{\Bernoulli}{\mathrm{Bernoulli}}
\newcommand{\Betadist}{\mathrm{Beta}}
\newcommand{\Uniform}{\mathrm{Uniform}}
\newcommand{\BG}{\mathrm{BernoulliGraph}}
\newcommand{\Categorical}{\mathrm{Categorical}}
\newcommand{\Multinomial}{\mathrm{Multinomial}}
\newcommand{\RDPG}{\mathrm{RDPG}}
\newcommand{\GRDPG}{\mathrm{GRDPG}}
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\theoremstyle{example}
\newtheorem{example}{Example}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\as}{\stackrel{\mathrm{a.s.}}{\to}}
\newcommand{\ind}{\stackrel{\mathrm{ind}}{\sim}}

@@

## Theoretical Framework

If we assume that the points are distributed as $X_i \ind \mathcal{N}(f(t_i | p), \Sigma(t_i | \phi))$, then we can write the (incomplete) log likelihood as:

$$\ell(p, \phi) = -\frac{1}{2} \sum_i |\Sigma(t_i | \phi)| - \frac{1}{2} \sum_i (x_i - f(t_i | p))^\top (\Sigma(t_i | \phi))^{-1} (x_i - f(t_i | p))$$

In the case $\Sigma(t_i | \phi) = \sigma^2 I$ and $f(t_i | p) = p^\top t_i$ where $t_i \in \mathbb{R}^{R+1}$ is the Bezier polynomial expansion of order $R$, then this becomes

$$\ell(p, \sigma^2) = -\frac{n d}{2} \log \sigma^2 - \frac{1}{2 \sigma^2} \|X - T p\|^2_F$$

Then given $t_1, ..., t_n$, the MLE of $p$ is again $(T^\top T)^{-1} T^\top X$, and the MLE of $\sigma^2$ is just $\frac{1}{n d} \|X - T \hat{p}\|^2_F$.
And given $p$, we can estimate each $t_i$ in the same way as before to maximize $\ell$ ($\sigma^2$ is unnecessary here).

### 2021-11-17 curve-fitting-methods/curve-fitting-scratch.tex
still working on typing up new material

>>>>

\hypertarget{theoretical-framework}{%
\subsection{Theoretical Framework}\label{theoretical-framework}}

If we assume that the points are distributed as
\(X_i \stackrel{\mathrm{ind}}{\sim}\mathcal{N}(f(t_i | p), \Sigma(t_i | \phi))\),
then we can write the (incomplete) log likelihood as:

\[\ell(p, \phi) = -\frac{1}{2} \sum_i |\Sigma(t_i | \phi)| - \frac{1}{2} \sum_i (x_i - f(t_i | p))^\top (\Sigma(t_i | \phi))^{-1} (x_i - f(t_i | p))\]

In the case \(\Sigma(t_i | \phi) = \sigma^2 I\) and
\(f(t_i | p) = p^\top t_i\) where \(t_i \in \mathbb{R}^{R+1}\) is the
Bezier polynomial expansion of order \(R\), then this becomes

\[\ell(p, \sigma^2) = -\frac{n d}{2} \log \sigma^2 - \frac{1}{2 \sigma^2} \|X - T p\|^2_F\]

Then given \(t_1, ..., t_n\), the MLE of \(p\) is again
\((T^\top T)^{-1} T^\top X\), and the MLE of \(\sigma^2\) is just
\(\frac{1}{n d} \|X - T \hat{p}\|^2_F\). And given \(p\), we can
estimate each \(t_i\) in the same way as before to maximize \(\ell\)
(\(\sigma^2\) is unnecessary here).
